Some examples in extended logic counting classes

### 2016-12-23 LICS/saluja.tex
Avance en sección de saluja.
Sección de Saluja primera versión.

@@

We see that many of the results in Saluja et. al. \cite{DBLP:journals/jcss/SalujaST95} for $\#\LL$ still apply in $\eqso(\LL)$ for a given fragment $\LL$:

\begin{theorem} \label{eqso-sigma-zero-in-fp}
	For every $\eqso(\loge{0})$ formula $\alpha$ over a signature $\R$, the function $f$ over $\R$ defined as $f(\enc(\A)) = \sem{\alpha}(\A)$ is in $\fp$.
\end{theorem}

\begin{theorem} \label{eqso-sigma-one-in-eqso-pi-one}
	For every $\eqso(\loge{1})$ formula $\alpha$ over a signature $\R$ there exists a $\eqso(\logu{1})$ formula $\beta$ over $\R$ such that $\sem{\alpha}(\A) = \sem{\beta}(\A)$ for every $\A\in\ostr[\R]$.
\end{theorem}

The {\em decision problem} associated to a function $f$ is defined by the language $L_f = \{\A \in \str \mid f(\A) > 0\}$.

\begin{theorem} \label{decisionptime}
	The decision problem associated to a function in $\eqso(\logex{1})$ is in \textsc{P}.
\end{theorem}

For a given pair of functions $f,g$, we define $f \dotminus g$ as follows:
\begin{eqnarray*}
	(f \dotminus g)(\A) =
	\begin{cases}
		f(\A)-g(\A), & \text{if }f(\A)>g(\A) \\
		0, & \text{if }f(\A) \leq g(\A).
	\end{cases}
\end{eqnarray*}
for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction} if for every pair of functions $f,g \in \F$, it holds that $f \dotminus g \in \F$.

\begin{theorem} \label{sub-pnp}
	If $\eqso(\loge{1})$ is closed under substraction, then {\sc P} = {\sc NP}.
\end{theorem}

\begin{theorem} \label{sigma1strict}
	$\eqso(\loge{1}) \subsetneq \eqso(\logex{1})$
\end{theorem}

For a given function $f$, we define $f \dotminus 1$ as follows:
\begin{eqnarray*}
	f \dotminus 1(\A) =
	\begin{cases}
		f(\A)-1, & \text{if }f(\A) > 0 \\
		0, & \text{if }f(\A) = 0.
	\end{cases}
\end{eqnarray*}
for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.

\begin{theorem} \label{sigmafo-minusone}
	$\eqso(\logex{1})$ is closed under substraction by one.
\end{theorem}

\begin{theorem} \label{dnf-pars}
	{\sc \#DNF} is hard for $\eqso(\loge{1})$ under parsimonious reductions. 
\end{theorem}

\begin{theorem} \label{nplusone-strict}
	$\U{1}$ with $n$ open first-order variables is properly contained in $\U{1}$ with $n+1$ open first-order variables for $n\in\N$.  
\end{theorem}

>>>>

%\begin{enumerate}
%	\item Introduce the classes $\eqso(\LL)$.
%	
%	\item Study the properties of $\eqso(\LL)$.
%	\begin{itemize}
%		\item If $\LL$ is closed under conjunction, then binary product can be defined in $\eqso(\LL)$.
%		
%		\item If $\LL$ is closed under XXX, then the first order product can be defined in $\eqso(\LL)$.
%		
%		\item Every formula in $\eqso(\LL)$ can be rewrite in \emph{sum normal form}.
%	\end{itemize}
%	
%	\item Introduce Saluja classes.
%	
%	\item Show the connection with Saluja classes.
%	\begin{itemize}
%		\item New classes are properly contained.
%		\item Old classes are properly contained/equivalent in new classes. 
%	\end{itemize}
%	
%	\item Study the complexity of new classes.
%	\begin{itemize}
%		\item $\eqso(\Sigma_0)$ is in FP.
%		\item $\eqso(\Sigma_1)$ is $\#P$ easy.
%		\item $\eqso(\Sigma_1)$ has free pass.
%	\end{itemize}
%	
%	\item Closure properties.
%	\begin{itemize}
%		\item $\eqso(\Sigma_1)$ is not closed under substraction.
%		\item $\eqso(\Sigma_1)$ is not closed under minus 1.
%	\end{itemize}
%	
%\end{enumerate}

Inspired by the connection between $\shp$ and $\#\fo$, the authors in~\cite{DBLP:journals/jcss/SalujaST95} studied subclasses of  $\#fo$ syntactical restricting the alternation quantifiers in boolean formulas defining the functions.
Specifically, they define the classes $\E{i}$ and $\U{i}$ where functions are defined by $\fo$-formulas in $\loge{i}$ and $\logu{i}$, respectively, for every $i \geq 0$. 
Interestingly, in~\cite{DBLP:journals/jcss/SalujaST95} they show that this function classes defined a finite hierarchy of the form:
\[
\E{0} \; \subsetneq \; \E{1} \; \subsetneq \; \U{1} \; \subsetneq \; \E{2} \; \subsetneq \; \U{2} \; = \; \#\fo 
\]
In this section, we study a natural extension of these classes in light of the $\eqso$ logic. Formally, we define the classes $\eqso(\loge{i})$ and $\eqso(\logu{i})$ where the boolean logic is restricted to $\loge{i}$ and $\logu{i}$, respectively. 
In the next result, we show that the prenex normal form hierarchy of $\eqso$ also defined a finite hierarchy similar than in~\cite{DBLP:journals/jcss/SalujaST95}.
\begin{proposition}
\begin{multline*}
\eqso(\loge{0}) \; \subsetneq \; \eqso(\loge{1}) \; \subsetneq \; \eqso(\logu{1}) \; \subsetneq \; \\ \eqso(\loge{2}) \; \subsetneq \; \eqso(\logu{2}) \; = \; \eqso(\fo)
\end{multline*}
\end{proposition}
A natural question until this point is what is the connection between $\E{i}$ and $\U{i}$ and their corresponding classes $\eqso(\loge{i})$ and $\eqso(\logu{i})$. Clearly, we have that $\E{i} \subseteq \eqso(\loge{i})$ and $\E{i} \subseteq \eqso(\logu{i})$.
The following result shows that this inclusion is strict only for the first levels of the hierarchy.
\begin{proposition}
	For $i \leq 1$, it holds that $\E{i} \subsetneq \eqso(\loge{i})$ and $\E{i} \subsetneq \eqso(\logu{i})$. Moreover, for $i \geq 2$ we have that $\E{i} = \eqso(\loge{i})$ and $\E{i} = \eqso(\logu{i})$.
\end{proposition}

Now, we study the complexity classes describe by this hierarchies. As the following result shows, $\eqso(\loge{0})$ defines only tractable counting functions and $\eqso(\loge{1})$ intractable counting functions but with an tractable decision problems. 
\begin{proposition}
All functions defined in $\eqso(\loge{0})$ and $\eqso(\loge{1})$ can be computed in $\fp$ and $\totp$, respectively. Furthermore, every function defined in $\eqso(\loge{1})$ has a FPTRAS.
\end{proposition}
Therefore, in terms of counting complexity, the $\eqso$-hierarchy behaves exactly the same as the $\#\fo$-hierarchy.

The next step is to study the closure properties of $\eqso$-hierarchy.
An advantage of the $\eqso$-hierarchy is that, by its language syntax, all the classes are closed under addition and first and second order sum. 
Furthermore, by Theorem~\ref{theo:binary-prod} one can easily see that all the hierarchy is also closed under binary product. 
Then a natural question is whether the hierarchy is closed under subtraction. Formally, for any pair of functions $f,g$, we define $f - g$ as the function such that $(f - g)(\A) = f(\A)-g(\A)$ whenever $f(\A)>g(\A)$ and $0$ otherwise.
As the next result shows, all classes in the $\eqso$-hierarchy is not closed under subtraction unless ${\sc P} = {\sc NP}$
\begin{theorem} \label{sub-pnp}
If $\eqso(\loge{i})$ or $\eqso(\logu{i})$ is closed under subtraction for $i > 0$, then {\sc P} = {\sc NP}.
\end{theorem}
\cristian{Martin, el resultado que tienes en el apendice se generaliza trivialmente para todas las clases ya que todas contienen la clase $\eqso(\loge{0})$.}

By the previous result, we know that functions in the $\eqso$ hierarchy are not closed under subtraction. Then, a natural restriction to this question is to ask whether these classes are closed under subtraction by one, namely, if $\CC$ is a class of functions and $f \in \CC$, is $f-1 \in \CC$ where $1$ is the constant function that outputs $1$ for every structure. 
We do not know $\eqso(\logex{1})$ is closed under subtraction by one. However, if we extend $\logex{1}$ with $\fo$ predicates.
\begin{theorem} \label{sigmafo-minusone}
	$\eqso(\logex{1})$ is closed under substraction by one.
\end{theorem}

@@

%We see that many of the results in Saluja et. al. \cite{DBLP:journals/jcss/SalujaST95} for $\#\LL$ still apply in $\eqso(\LL)$ for a given fragment $\LL$:
%
%\begin{theorem} \label{eqso-sigma-zero-in-fp}
%	For every $\eqso(\loge{0})$ formula $\alpha$ over a signature $\R$, the function $f$ over $\R$ defined as $f(\enc(\A)) = \sem{\alpha}(\A)$ is in $\fp$.
%\end{theorem}
%
%\begin{theorem} \label{eqso-sigma-one-in-eqso-pi-one}
%	For every $\eqso(\loge{1})$ formula $\alpha$ over a signature $\R$ there exists a $\eqso(\logu{1})$ formula $\beta$ over $\R$ such that $\sem{\alpha}(\A) = \sem{\beta}(\A)$ for every $\A\in\ostr[\R]$.
%\end{theorem}
%
%
%The {\em decision problem} associated to a function $f$ is defined by the language $L_f = \{\A \in \str \mid f(\A) > 0\}$.
%
%\begin{theorem} \label{decisionptime}
%	The decision problem associated to a function in $\eqso(\logex{1})$ is in \textsc{P}.
%\end{theorem}

%For a given pair of functions $f,g$, we define $f \dotminus g$ as follows:
%\begin{eqnarray*}
%	(f \dotminus g)(\A) =
%	\begin{cases}
%		f(\A)-g(\A), & \text{if }f(\A)>g(\A) \\
%		0, & \text{if }f(\A) \leq g(\A).
%	\end{cases}
%\end{eqnarray*}
%for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction} if for every pair of functions $f,g \in \F$, it holds that $f \dotminus g \in \F$.
%
%\begin{theorem} \label{sub-pnp}
%	If $\eqso(\loge{1})$ is closed under substraction, then {\sc P} = {\sc NP}.
%\end{theorem}
%
%\begin{theorem} \label{sigma1strict}
%	$\eqso(\loge{1}) \subsetneq \eqso(\logex{1})$
%\end{theorem}
%
%For a given function $f$, we define $f \dotminus 1$ as follows:
%\begin{eqnarray*}
%	f \dotminus 1(\A) =
%	\begin{cases}
%		f(\A)-1, & \text{if }f(\A) > 0 \\
%		0, & \text{if }f(\A) = 0.
%	\end{cases}
%\end{eqnarray*}
%for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
%
%\begin{theorem} \label{sigmafo-minusone}
%	$\eqso(\logex{1})$ is closed under substraction by one.
%\end{theorem}
%
%\begin{theorem} \label{dnf-pars}
%	{\sc \#DNF} is hard for $\eqso(\loge{1})$ under parsimonious reductions. 
%\end{theorem}
%
%\begin{theorem} \label{nplusone-strict}
%	$\U{1}$ with $n$ open first-order variables is properly contained in $\U{1}$ with $n+1$ open first-order variables for $n\in\N$.  
%\end{theorem}

### 2016-12-23 LICS/syntactic.tex
Avance en sección de saluja.
Sección de Saluja primera versión.

\subsection{Saluja Classes}

\subsection{Horn Counting Classes}

>>>>

In this section we study the fragment of $\eqso(\LL)$ when $\LL$ is a boolean logic contained in $\fo$.

We say that a formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\Sigma$-prenex normal form} if $\alpha$ is of the form:
\[
\alpha := \sa{\bar{X}} \sa{\bar{x}} \varphi(\bar{X}, \bar{x})
\]
where $\bar{X}$ and $\bar{x}$ is a sequence of $0$ or more second and first order variables, respectively.
In particular, note that a boolean formula is always in $\Sigma$-prenex normal form.
Then we say that a formula $\alpha$ is in \emph{sum normal form} if $\alpha$ is of the form $c + \Sigma_{i=1}^n \alpha_i$ where $c$ is a non-negative constant and each $\alpha_i$ is in $\Sigma$-prenex normal form.
Next we show that each formula in $\eqso(\LL)$ can be converted in \emph{sum normal form} but not always in prenex normal form.

\begin{theorem}
	Every formula in $\eqso(\LL)$ can be rewritten in \emph{sum normal form}. Furthermore, there exists a language $\LL$ and a formula $\alpha$ such that there does not exist a formula in $\Sigma$-prenex normal form equivalent to $\alpha$.
\end{theorem}

An interesting question here is whether the multiplicative operators in $\qso$ can be defined in $\eqso(\LL)$. As the following result shows, if $\LL$ is closed under conjunction, then the binary product can be defined in  $\eqso(\LL)$.

\begin{theorem}\label{theo:binary-prod}
	If $\LL$ is closed under conjunction, then binary product can be defined in $\eqso(\LL)$.
\end{theorem}

\subsection{Prenex-normal form classes}
\input{saluja}

\subsection{Horn Counting Classes}
\input{horn-counting}
